Learning Causally-Generated Stationary Time Series

### 2017-03-15 doc/paper.tex
First commit of source, still broken
Typos
Added draft of section on choice of hyperparameters

% Break in equations...
\allowdisplaybreaks

@@

% Operators
\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\max}{\operatorname{max}}
\newcommand{\argmax}{\operatorname{argmax}}
\newcommand{\argmin}{\operatorname{argmin}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\sign}{\operatorname{sign}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\card}{\#}
\newcommand{\comp}{\circ}
\newcommand{\had}{\circ}
\newcommand{\chol}{\operatorname{chol}}
\newcommand{\ind}{1}
\newcommand{\KL}{\operatorname{D}_{\text{KL}}}

% Special math commands
\renewcommand{\ss}[1]{_\mathit{#1}}       % Subscripts without spacing
\newcommand{\id}[1]{\, \mathrm{d} #1}     % Straight 'd' in integral
\newcommand{\sce}{\text{\sc{e}}}          % Scientific notation
\newcommand{\cond}{\, | \,}               % Conditioning
\renewcommand{\ll}{\left}
\newcommand{\rr}{\right}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\phan}[1]{\hphantom{#1\;}}

@@

@@

\begin{figure*}[t]
    \centering
    GPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_gpcm.pdf}
    CGPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a noisy random sample from the GPCM. The learned sample consists of 300 data points. We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
    \label{fig:toy_acausal_sample}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    GPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_gpcm.pdf}
    CGPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a noiseless random sample from the CGPCM. The learned sample consists of 300 data points. We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
    \label{fig:toy_causal_sample}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
        GPCM & CGPCM
    \end{tabularx}
    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
        $k_{f\cond h}$ & $h$ & $k_{f\cond h}$ & $h$
    \end{tabularx}
    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_gpcm.pdf}
    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a HRIR convolved with white noise. The learned sample consists of 500 data points. We used 101 evenly-spaced inducing points for $h$ and 150 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
    \label{fig:hrir}
\end{figure*}

>>>>

% Break in equations...
% \allowdisplaybreaks

@@

% Operators
\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\max}{\operatorname{max}}
\newcommand{\atan}{\operatorname{atan}}
\newcommand{\argmax}{\operatorname{argmax}}
\newcommand{\argmin}{\operatorname{argmin}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\sign}{\operatorname{sign}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\card}{\#}
\newcommand{\comp}{\circ}
\newcommand{\had}{\circ}
\newcommand{\chol}{\operatorname{chol}}
\newcommand{\ind}{1}
\newcommand{\KL}{\operatorname{D}_{\text{KL}}}

% Special math commands
% \renewcommand{\ss}[1]{_\mathit{#1}}       % Subscripts without spacing
\renewcommand{\ss}[1]{_\text{#1}}         % Subscripts in text
\newcommand{\id}[1]{\, \mathrm{d} #1}     % Straight 'd' in integral
\newcommand{\sce}{\text{\sc{e}}}          % Scientific notation
\newcommand{\cond}{\, | \,}               % Conditioning
\renewcommand{\ll}{\left}
\newcommand{\rr}{\right}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\phan}[1]{\hphantom{#1\;}}

@@

\subsection{Choice of Hyperparameters}
Following \citet{Pavliotis:2012:Probabilistic_Perspective}, let the \textit{correlation time} $\tau$, or alternatively the \textit{length scale}, of a stationary process with kernel $k(r)$ be
\begin{align*}
    \tau = \frac{1}{k(0)}\int_0 k(r) \id{r}.
\end{align*}

Let $\tau\ss{ac}$ and $P\ss{ac}$ and $\tau\ss{c}$ and $P\ss{c}$ be respectively the correlation time and power of the process $f$ for the GPCM and CGPCM respectively. Then Appendix XX shows that $\tau_w=\sqrt{\pi/8 \alpha}$ and
\begin{align*}
    P\ss{ac} &= \sigma_{f,\text{ac}}^2\sqrt{\frac{\pi}{2\alpha\ss{ac}}},
    &\tau\ss{ac} &= \sqrt{\frac{\pi}{2(\alpha\ss{ac} + 2 \gamma\ss{ac})}}  \\
    P\ss{c} &= \sigma_{f,\text{c}}^2\sqrt{\frac{\pi}{8\alpha\ss{c}}},
    &\tau\ss{c} &= \sqrt{\frac{2}{\pi(\alpha\ss{c} + 2 \gamma\ss{c})}} \\
    &&&\phan{=}\atan\ll( \sqrt{\frac{\alpha\ss{c} + 2 \gamma\ss{c}}{\alpha\ss{c}}} \rr). \\
\end{align*}
It follows that $P\ss{c}=\frac{1}{2}P\ss{ac}$ and $\tau\ss{c}\approx \tau\ss{ac}$ if $\gamma\ss{c} \gg \alpha\ss{c}$, $\alpha\ss{c}=\alpha\ss{ac}$, $\gamma\ss{c}=\gamma\ss{ac}$, and $\sigma_{f,\text{c}}^2=\sigma_{f,\text{ac}}^2$. To allow for fair comparison, we require that $P\ss{ac}=P\ss{c}$ and $\tau\ss{ac}=\tau\ss{c}$. If we further choose $P\ss{ac}=P\ss{c}=1$ and $\sigma_{f,\text{ac}}^2=\sigma_{f,\text{c}}^2$, then we find that $\alpha\ss{c}=\frac{1}{4}\alpha\ss{ac}$, $\gamma\ss{c}=\frac{3}{8}\alpha\ss{ac}+ \gamma\ss{ac}$, and $\sigma_{f,\text{ac}}^2 = \sigma_{f,\text{c}}^2 = \sqrt{2\alpha\ss{ac}/\pi}$.

To determine $\alpha\ss{ac}$ and $\gamma\ss{ac}$, we specify the correlation time $\tau_w$ of the window $w$ and the correlation time $\tau_f$ of the process $f$. Then Appendix XX shows that
\begin{align*}
    \alpha\ss{ac} &= \frac{\pi}{2} \cdot \frac{1}{2 \tau_w^2},
    &\gamma\ss{ac} &= \frac{\pi}{2} \cdot \frac{1}{2 \tau^2_f} - \frac{1}{2} \alpha\ss{ac}.
\end{align*}
Since $\alpha\ss{ac} > 0$ and $\gamma\ss{ac} > 0$, it must hold that $\sqrt{2} \tau_w > \tau_f > 0$.

Furthermore, for the GPCM, we let $t_u$ be evenly spaced from $-3\tau_w$ to $3 \tau_w$. For the CGPCM, we let $t_u$ be evenly spaced from zero to $6 \tau_w$, shifted two inter-point spacings $\Delta t_u$ to the left to account for derivatives of $h$ at zero.

Finally, we let $t_z$ be evenly spaced on the domain of interest, and choose $\omega$ such that $s$'s correlation time $\tau_s$ equals twice $t_z$'s inter-point spacing $\Delta t_z$: $\omega = \pi / 8 \Delta t_z^2$.

@@

\begin{figure*}[t]
    \centering
    GPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_gpcm.pdf}
    CGPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a noisy random sample from the GPCM. The learned sample consists of 300 data points. We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
    \label{fig:toy_acausal_sample}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    GPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_gpcm.pdf}
    CGPCM
    \begin{tabularx}{\linewidth}{
            >{\centering}>{\hsize=.43\hsize}X  % ??
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
            >{\centering}>{\hsize=.2\hsize}X
        }
        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
    \end{tabularx}
    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a noiseless random sample from the CGPCM. The learned sample consists of 300 data points. We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
    \label{fig:toy_causal_sample}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
        GPCM & CGPCM
    \end{tabularx}
    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
        $k_{f\cond h}$ & $h$ & $k_{f\cond h}$ & $h$
    \end{tabularx}
    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_gpcm.pdf}
    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_cgpcm.pdf}
    \caption{Predictions of the GPCM and CGPCM after learning a HRIR convolved with white noise. The learned sample consists of 500 data points. We used 101 evenly-spaced inducing points for $h$ and 150 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure. All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed. All signals are normalised to unity power. {\color{red}Red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation. For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
    \label{fig:hrir}
\end{figure*}
